Towards Practical First-Order Model Counting

### 2024-03-04 doc/paper.tex
The beginnings of some data-plotting code
Writing about smoothing

\paragraph{Changes to the compilation process.}
\begin{itemize}
  \item We make the FCGs satisfy the following property. For every domain
        $\Delta$, for every directed path $P$ without Ref nodes, the number of
        atom counting or (generalised) domain recursion or independent partial
        grounding operations on $\Delta$ on $P$ is at most one.
  \item when checking whether recursion is possible between a source and a
        target formula, the target formula (maybe the source formula as well?)
        must not include a domain as well as its subdomain. This might be
        equivalent to the formula being `unshattered'. Such formulas could be
        seen as transitory, and allowing such recursive calls leads to FCGs that
        I cannot make sense of.
  \item During smoothing, when unit clauses (a.k.a.\ variables) are propagated
        in the opposite direction of the FCG arcs (i.e., `upwards'), when
        visiting a \texttt{Ref} node, these clauses are translated using the
        domain map of the \texttt{Ref} node. For example, we might have a
        function $g$ that depends on $m$ (i.e., domain $a^{\top}$), which is
        called as $g(n)$ (i.e., on domain $a$). When smoothing that function
        call, we want to translate all mentions of $a^{\top}$ to $a$.
  \item Implemented an if-then-else arithmetic operation as the interpretation
        of the new domain recursion node (i.e., there is one branch for when the
        domain is empty and another branch for any other domain size).
  \item Smoothing for domain recursion now works the same way as smoothing for
        disjunction: both sides must have the same atoms. Almost all of
        \cref{sec:simplifying} is now outdated because smoothing makes the
        ``tautology insertion'' unnecessary.
  \item \Cref{sec:main,sec:identifying} might also be outdated: there is
        probably no need to search for base cases given the new version of
        domain recursion.
\end{itemize}

@@

\paragraph{Contributions.}
\begin{itemize}
  \item \Cref{sec:identifying}
  \item \Cref{sec:simplifying}
  \item Converting the recursive equations into a C++ program, which can then be
        compiled and executed to obtain numerical values (see \cref{sec:cpp}).
  \item Support for infinite precision integers using the GNU Multiple Precision
        Arithmetic Library.
  \item The experiments are (going to be) more comprehensive than any other
        experimental study performed on (W)FOMC algorithms.
\end{itemize}

@@

>>>>

\paragraph{Changes to the compilation process.}
\begin{itemize}
  \item We make the FCGs satisfy the following property. For every domain
        $\Delta$, for every directed path $P$ without Ref nodes, the number of
        atom counting or (generalised) domain recursion or independent partial
        grounding operations on $\Delta$ on $P$ is at most one.
  \item when checking whether recursion is possible between a source and a
        target formula, the target formula (maybe the source formula as well?)
        must not include a domain as well as its subdomain. This might be
        equivalent to the formula being `unshattered'. Such formulas could be
        seen as transitory, and allowing such recursive calls leads to FCGs that
        I cannot make sense of.
  \item Implemented an if-then-else arithmetic operation as the interpretation
        of the new domain recursion node (i.e., there is one branch for when the
        domain is empty and another branch for any other domain size).
  \item \Cref{sec:main,sec:identifying} might also be outdated: there is
        probably no need to search for base cases given the new version of
        domain recursion.
\end{itemize}

@@

\paragraph{Contributions.}
\begin{itemize}
  \item \Cref{sec:identifying}
  \item \Cref{sec:simplifying}
  \item Converting the recursive equations into a C++ program, which can then be
        compiled and executed to obtain numerical values (see \cref{sec:cpp}).
  \item Support for infinite precision integers using the GNU Multiple Precision
        Arithmetic Library.
  \item A new set of benchmarks.
  \item The experiments for Crane and ForcLift are (going to be) more
        comprehensive than any other experimental study performed on (W)FOMC
        algorithms.
  \item Some experiments vs FastWFOMC
  \item A proof of correctness
  \item A proof of completeness
\end{itemize}

@@

\section{Smoothing}

Goal of smoothing: whenever rules such as unit propagation or
inclusion-exclusion (maybe just these two?) eliminate the consideration of some
ground atoms, smoothing nodes should be inserted `at the same level' so that
these ground atoms are still considered when counting. Note that conjunction
nodes are irrelevant here because multiplication is associative. `At the same
level' means that, e.g., if some ground atoms were eliminated from consideration
before/after performing independent partial grounding, then the smoothing node
should also appear before/after the independent partial grounding node.

\todo[inline]{ Can I prove that the additions to smoothing outlined in the rest
  of this section `do the right thing'? }

\subsection{Stage 1: Propagating Unit Clauses `Upwards'}

\begin{description}
  \item[\texttt{Ref}:] during smoothing, when unit clauses (a.k.a.\ variables)
        are propagated in the opposite direction of the FCG arcs (i.e.,
        `upwards'), when visiting a \texttt{Ref} node, these clauses are
        translated using the domain map of the \texttt{Ref} node. For example,
        if the domain map include $\Delta \mapsto \Delta'$, and the unit clause
        mentions $\Delta$, then replace it by $\Delta'$.
  \item[Constraint removal:] do reverse constraint removal. Assuming that
        domain $\Delta$ with constraints $X \ne c$ (for some constant
        $c \in \Delta$) were replaced with domain $\Delta'$, replace each
        $\forall X \in \Delta'\text{. }\phi(X)$ with
        $\forall X \in \Delta\text{. }X \ne c \Rightarrow \phi(X)$.
  \item[Domain recursion:] nothing special needs to be done (just propagate the
        union of the two sets of unit clauses upwards), assuming that $P(c)$ and
        $\forall X \in \Delta\text{. }X \ne c \Rightarrow P(X)$ are
        automatically combined into $\forall X \in \Delta\text{. }P(X)$.
\end{description}

\subsection{Stage 2: Adding Smoothing Nodes}

\begin{description}
  \item[\texttt{Ref}:] nothing.
  \item[Constraint removal:] nothing.
  \item[Domain recursion:] Smoothing for domain recursion now works the same way
        as smoothing for disjunction: both sides must have the same atoms EXCEPT
        the `base case' side need not include unit clauses that mention the
        constant introduced by this domain recursion node or its domain (since
        this branch assumes that this domain is empty). NOTE: As a consequence,
        almost all of \cref{sec:simplifying} is now outdated because smoothing
        makes the ``tautology insertion'' unnecessary.
\end{description}
